### 2022-08-28 hw1-1/ashish/a.svm_27082022.Rmd
added 2.1 part 1,2

---
title: "SVM modelling iterations Ashish"
output: github_document
---

```{r}
library(kernlab)
library(ggplot2)
```

### Read Data

```{r}
org_cc_data <- read.table(file = '../data 2.2/credit_card_data-headers.txt', sep = "\t", header = TRUE)
dim(org_cc_data)
names(org_cc_data)
```

### v0: Vanilla Model and Accuracy

```{r}
model_v0 = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = 10)
#model_v0

acc_func <- function(model) {
   pred_all <- predict(model,org_cc_data[,1:10])
   print (paste("Overall Acc:", round(sum(pred_all == org_cc_data[,11]) * 100 / nrow(org_cc_data),4)))

   pred_1 <- predict(model,org_cc_data[org_cc_data$R1 == 1,1:10])
   print (paste("Acc in 1's:", round(sum(pred_1 == org_cc_data[org_cc_data$R1 == 1,11]) * 100 / nrow(org_cc_data[org_cc_data$R1 == 1,1:10]),4)))

   pred_0 <- predict(model,org_cc_data[org_cc_data$R1 == 0,1:10])
   print (paste("Acc in 0's:", round(sum(pred_0 == org_cc_data[org_cc_data$R1 == 0,11]) * 100 / nrow(org_cc_data[org_cc_data$R1 == 0,1:10]),4)))
}

acc_func(model_v0)
```

### v1: Optimise C: Find best window for linear search

```{r}
C_values <- c(0.0001,0.001,0.0015,0.002,0.005,0.01,0.03)

for (C_i in C_values) {
  print (paste("For C = ",C_i))
  acc_func(ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = C_i))
  print("")
}
```

### v2: vanilla but capping A15

```{r}
hist(org_cc_data$A15, freq = FALSE, breaks = 100)
hist(log(org_cc_data$A15), freq = FALSE, breaks = 100)
#lines(density(x))

cc_data1 = org_cc_data
cc_data1$A15 = log(0.001 + cc_data1$A15)
acc_func(ksvm(x=as.matrix(cc_data1[,1:10]), y=cc_data1[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = 0.007))
```
### Final model and equation

```{r}
modelf = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = 0.0015)
acc_func(modelf)

a <- colSums(modelf@xmatrix[[1]] * modelf@coef[[1]])
a0 <- -modelf@b
a0
```

### Non Linear models: Radial Basis Kernel
```{r}
C_values <- c(0.01,1,10,50,100,1000,2000)

for (C_i in C_values) {
  print (paste("For C = ",C_i))
  modeli = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "rbfdot", C = C_i)
  acc_func(modeli)
  print(modeli@nSV)
}
```

### Non Linear models: Laplace Kernel
```{r}
C_values <- c(0.01,1,10,50,100,1000,2000)

for (C_i in C_values) {
  print (paste("For C = ",C_i))
  modeli = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "laplacedot", C = C_i)
  acc_func(modeli)
  print(modeli@nSV)
}
```