### 2022-08-28 hw1-1/eda_27082022.Rmd

---
title: "eda"
output: github_document
---

```{r}
#imports
library(cowplot)
library(ggplot2)
library(reshape2)
```

### Credit Card Data Intro EDA

1.  Read data

```{bash}
head -5 ./data\ 2.2/credit_card_data.txt
```

```{r}
org_cc_data <- read.table(file = './data 2.2/credit_card_data-headers.txt', sep = "\t", header = TRUE)
dim(org_cc_data)
class(org_cc_data)
org_cc_data
```

2.  Summary of Data

```{r}
summary(org_cc_data)
#class(summary(org_cc_data))
```

**A1,A9,A10,A12 are binary basis min/max values, rest are continuous**
**For target variable mean is ~45% ==> variable is not imbalanced**

3.  Distribution of Data

```{r}
my_plots <- lapply(names(org_cc_data), function(var_x){
  p <- 
    ggplot(org_cc_data) +
    aes_string(var_x)

  if(var_x %in% list("A1","A9","A10","A12","R1")) {
    p <- p + geom_bar()

  } else {
    p <- p + geom_density()
  }

})

plot_grid(plotlist = my_plots)
#class(summary(org_cc_data))
```
**A15 seems highly skewed towards right side and A11 seems slightly skewed**

4.  Correlation Heat Map
```{r}
cormat <- round(cor(org_cc_data),2)
cormat[upper.tri(cormat)] <- NA
melted_cormat <- melt(cormat)
# plotting the correlation heatmap
library(ggplot2)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2,
                                   fill=value)) +
geom_tile() +
geom_text(aes(Var1, Var2, label = value),
          color = "white", size = 3)
```
If |cor| > 0.3 ==> Significant (Let)
R1: A8,A9,A10,A11
A11: A8,A9,A10
A10: A9
A9: A8
A8: A2