### 2022-08-28 hw1-1/abhinav/knn.Rmd
KNN classification script

---
title: "R Notebook"
output: github_document
---

```{r}
#imports
library(kknn)
```

```{r}
knn_data<-read.table("./hw1-1/data 2.2/credit_card_data-headers.txt",sep='\t',header=TRUE)
dim(knn_data)
```

```{r}
# Plain version with kernel as optimised 
knn_pred<-vector(mode="list",length = dim(knn_data)[1])
pred_table<-data.frame(k_val=numeric(0),acc_val=double(0))
for (k0 in seq(1,nrow(knn_data),by=10)){
  for (i in 1:nrow(knn_data)){
    knn_model_v0<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="optimal",scale = TRUE)
    knn_pred[i]<-as.integer(round(fitted(knn_model_v0),digits = 0)==knn_data$R1[i]) 
  }
  percent_accuracy<-round(as.double(sum(unlist(knn_pred)))/nrow(knn_data),digits =2)
  pred_table[nrow(pred_table)+1,]=c(k0,percent_accuracy)
}
```

```{r}
# need to iterate for lower values of K
plot(pred_table)
```

```{r}
knn_pred<-vector(mode="list",length = dim(knn_data)[1])
pred_table<-data.frame(k_val=numeric(0),acc_val=double(0))
for (k0 in seq(1,25,by=3)){
  for (i in 1:nrow(knn_data)){
    knn_model_v0<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="optimal",scale = TRUE)
    knn_pred[i]<-as.integer(round(fitted(knn_model_v0),digits = 0)==knn_data$R1[i]) 
  }
  percent_accuracy<-round(as.double(sum(unlist(knn_pred)))/nrow(knn_data),digits =2)
  pred_table[nrow(pred_table)+1,]=c(k0,percent_accuracy)
}
```