### 2022-08-31 hw1-1/anshit/HW1.Rmd

# Assignment 1

### Question 2.2

*Exploratory Data Analysis*

```{r}
library(cowplot)
library(ggplot2)
library(reshape2)
library(kernlab)
```

```{r}
data <- read.table('../data 2.2/credit_card_data-headers.txt', header = TRUE, sep = "\t", dec = ".")
data
```

```{r}
summary(data)
```

```{r}
my_plots = lapply(colnames(data), function(x){
  p = ggplot(data) + aes_string(x)

  if(x %in% list("A1", "A9", "A10", "A12", "R1")){
    p = p + geom_bar()
  } else {
    p = p + geom_density()
  }
})

plot_grid(plotlist = my_plots)
```

**Correlation Matrix**

```{r}
cor_mat = round(cor(data), 2)
cor_mat[upper.tri(cor_mat)] = NA
melted_mat = melt(cor_mat)
ggplot(data = melted_mat, aes(x=Var1, y=Var2, fill=value)) + geom_tile() + 
  geom_text(aes(Var1, Var2, label=value), color="white", size=3)
```
*Modelling*

```{r}
# call ksvm.  Vanilladot is a simple linear kernel.
model <- ksvm(as.matrix(data[,1:10]),data[,11],type="C-svc",kernel="vanilladot",C=0.015,scaled=TRUE)
model
```

```{r}
#predict using model
ypred <- predict(model, as.matrix(data[,1:10]))
table(data[,11], ypred)
```

```{r}
### Accuracy function: Overall and in each class
acc_func <- function(model) {
   pred_all <- predict(model,org_cc_data[,1:10])
   print (paste("Overall Acc:", round(sum(pred_all == org_cc_data[,11]) * 100 / nrow(org_cc_data),4)))

   pred_1 <- predict(model,org_cc_data[org_cc_data$R1 == 1,1:10])
   print (paste("Acc in 1's:", round(sum(pred_1 == org_cc_data[org_cc_data$R1 == 1,11]) * 100 / nrow(org_cc_data[org_cc_data$R1 == 1,1:10]),4)))

   pred_0 <- predict(model,org_cc_data[org_cc_data$R1 == 0,1:10])
   print (paste("Acc in 0's:", round(sum(pred_0 == org_cc_data[org_cc_data$R1 == 0,11]) * 100 / nrow(org_cc_data[org_cc_data$R1 == 0,1:10]),4)))

   print(paste("#Support Vectors",model_v0@nSV))
}
```

```{r}
model_v0 = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = 10)

acc_func(model_v0)
```

```{r}
C_values <- c(0.0001,0.001,0.0015,0.002,0.005,0.01,0.03) #Range identified with hit and trial

for (C_i in C_values) {
  print (paste("For C = ",C_i))
  acc_func(ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = C_i))
  print("")
}
```

**Hypertuning Parameters**

```{r}
#Variables not seeming significant -> A1, A12, A14, A15

new_parameters = data[, c('A1', 'A2', 'A3', 'A8', 'A9', 'A10', 'A11', 'A12', 'A14', 'A15', 'R1')]
head(new_parameters)

```

```{r}
column_range = 1:10
x_train = new_parameters[,column_range]

model_v1 = ksvm(x=as.matrix(x_train), y=new_parameters[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C =0.0015)

### Accuracy function: Overall and in each class
acc_func <- function(model) {
   pred_all <- predict(model, x_train)
   print (paste("Overall Acc:", round(sum(pred_all == new_parameters[,11]) * 100 / nrow(new_parameters),4)))

   pred_1 <- predict(model,new_parameters[new_parameters$R1 == 1,column_range])
   print (paste("Acc in 1's:", round(sum(pred_1 == new_parameters[new_parameters$R1 == 1,11]) * 100 / nrow(new_parameters[new_parameters$R1 == 1, column_range]),4)))

   pred_0 <- predict(model,new_parameters[new_parameters$R1 == 0,column_range])
   print (paste("Acc in 0's:", round(sum(pred_0 == new_parameters[new_parameters$R1 == 0,11]) * 100 / nrow(new_parameters[new_parameters$R1 == 0, column_range]),4)))

   print(paste("#Support Vectors",model_v1@nSV))
}

acc_func(model_v1)
```