### 2022-09-06 hw2/ashish/a.0_kfold_knnn.Rmd
ques 3.1 complete

---
title: "ques 3.1"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, include=FALSE}
setwd("~/data_projects/fall22_hw/isye6501_aaa/hw2/ashish")

set.seed(520)

#lib imports
library(cowplot)
library(ggplot2)
library(reshape2)
library(kknn)
library(pROC)
```

### Read Data and repeat vanilla knnn model from last time

```{r}
org_cc_data <- read.table(file = '../data 3.1/credit_card_data-headers.txt', sep = "\t", header = TRUE)
dim(org_cc_data)
```

```{r}
#predict ith row of dataframe
predcit_i = function(i =1, k0 = 5){
  model = kknn(R1~.,train = org_cc_data[-i,],test = org_cc_data[i,],k=k0,kernel="rectangular",scale = TRUE)
  return (fitted(model))
  #return (as.integer(fitted(model) > threshold))
}

#predictions for a given k
predcit_k = function(k0 = 5){
  predictions = lapply(seq(1,nrow(org_cc_data)),predcit_i, k0)
  return (predictions)
}

#precit probabilities for specific k
pred_prob <- predcit_k(k0=5)

# decide best threshold
best_threshold = function(pred_prob, df_test=org_cc_data){
  myroc <- roc(df_test$R1,as.numeric(unlist(pred_prob)), smooth=FALSE)
  print(paste("AUC is",auc(myroc)))
  #plot(myroc, auc.polygon=TRUE)
  return (coords(myroc,x="best")["threshold"][[1]])
}

ti = best_threshold(pred_prob)
print (ti)
#Accuracy for best threshold
acc_func = function(y_act, pred_probs, thresh = 0.5){
  y_pred = lapply(pred_probs, function(x) if(x>=thresh) 1 else 0)
  y_pred = as.numeric(unlist(y_pred))
  acc_overall = sum(y_act == y_pred)/length(y_act)
  acc_overall = round(acc_overall,4)* 100
  return (acc_overall)
}

acc_func(y_act = org_cc_data$R1, pred_probs = pred_prob, thresh = ti)
acc_func(y_act = org_cc_data$R1, pred_probs = pred_prob, thresh = 0.9)
```

```{r}
master_kknn_func = function(y_act = org_cc_data$R1, ki = 10){
  pred_prob <- predcit_k(k0=ki)
  ti = best_threshold(pred_prob)
  print (paste("Best threshold",ti))
  return (acc_func(y_act = y_act, pred_probs = pred_prob, thresh = ti))
}

master_kknn_func(ki = 5)
master_kknn_func(ki = 7)
master_kknn_func(ki = 10)
```

## Cross validation for kknn

-   Split data into 5 parts

    -   Keep first 4 for train, last 1 for validation

    -   On the train, obtain k values of nearest neighbors for max auc

    -   get acc on train and test

```{r}
#Split org_data into 5 parts
df_copy = org_cc_data
df_copy$cross_validation_i <- sample(5, size = nrow(df_copy), replace = TRUE)

aggregate(df_copy$R1,list(df_copy$cross_validation_i), FUN = mean)
```

```{r}
#Function to train model and return prediction on test
train_test_i = function(df_train,df_test, k0 = 5){
  f = R1~A1+A2+A3+A8+A9+A10+A11+A12+A14+A15
  model = kknn(f,train = df_train,test = df_test,k=k0,kernel="rectangular",scale = TRUE)
  return (fitted(model))
}

#Train test for one group
cv_function_i = function(df= df_copy, test_index_choice,k_choice){
  print (paste("test index for CV Fold",test_index_choice))
  #split data into train and test
  mask_test_i = df$cross_validation_i == test_index_choice
  df_train = df[!mask_test_i,]
  df_test = df[mask_test_i,]
  print (paste("Train,Test Size =",nrow(df_train),nrow(df_test)))
  #train and get probaility predictions on test model
  test_probabilities = train_test_i(df_train,df_test, k0 = k_choice)

  #get best threshold for test probabilities
  ti = best_threshold(test_probabilities, df_test = df_test)
  if (length(ti) >1){
    ti = ti[[1]] #select 1st when there are multiple optimal thresholds
  }
  print (paste("Best threshold is",ti))
  #get test accuracy
  y_act = df_test$R1
  #print (paste("Length of Test Probabilities",nrow(df_test),length(test_probabilities)))
  return (acc_func(y_act = y_act, pred_probs = test_probabilities, thresh = ti))
}

#CV function: pass k, return test accuracy for all folds
cv_function = function(df = df_copy,k_choice = 5){
  cat(paste("\nFor k value",k_choice))
  cv_accuracy = lapply(seq(1,5), function(x) cv_function_i(df=df_copy,x,k_choice))
}
```

```{r}
cv_accuracy = cv_function(df = df_copy,k_choice = 5)
mean(unlist(cv_accuracy))
```

## Iterate on different models and use CV for best model

```{r, include=FALSE}
#Iterate over k and get mean test
acc_df <- data.frame(matrix(ncol = 0, nrow = length(seq(3,71,2))))
acc_df$k_kknn = seq(3,71,2)
mean_cv_acc_k = lapply(seq(3,71,2), function(x) mean(unlist(cv_function(df = df_copy,x))))
min_cv_acc_k = lapply(seq(3,71,2), function(x) min(unlist(cv_function(df = df_copy,x))))
acc_df$mean_cv_accuracy = unlist(mean_cv_acc_k)
acc_df$min_cv_accuracy = unlist(min_cv_acc_k)
#mean(unlist(cv_accuracy))
```

```{r}
acc_df
```

```{r}
ggplot(data = acc_df, aes(x = k_kknn, y = mean_cv_accuracy)) +
  geom_line() +geom_point() +
  geom_vline(xintercept = acc_df[which.max(acc_df$mean_cv_accuracy),1],
             color ="magenta", size = 1.5, alpha = 0.5) +
  ggtitle(paste("max accuracy for k =",acc_df[which.max(acc_df$mean_cv_accuracy),1]))

print (paste("max accuracy for k =",acc_df[which.max(acc_df$mean_cv_accuracy),1]))
```

## Ques 3.1: part 2

#split data into 70% train, 15% validation and 15% test

```{r}
#Split org_data into 5 parts
df_copy2 = org_cc_data
df_copy2$split_i <- sample(100, size = nrow(df_copy), replace = TRUE)

df_copy2$split = "train"
df_copy2$split[70 < df_copy2$split_i] = "valid"
df_copy2$split[85 < df_copy2$split_i] = "test"

summary_df = aggregate(df_copy2$R1,list(df_copy2$split), 
          FUN = function(x) c(mean = mean(x), proportion = length(x)/nrow(df_copy2)))
print.data.frame(summary_df)
```

Find optimal k using training on train and validation

```{r}
df_train = df_copy2[df_copy2$split=="train",]
dim(df_train)
df_valid = df_copy2[df_copy2$split=="valid",]
dim(df_valid)
df_test= df_copy2[df_copy2$split=="test",]
dim(df_test)
```

```{r}
train_valid_func = function(df_train, df_test, k_choice){
  print (paste("For k=",k_choice))
  #split data into train and test
  print (paste("Train,Test Size =",nrow(df_train),nrow(df_test)))
  #train and get probaility predictions on test model
  test_probabilities = train_test_i(df_train,df_test, k0 = k_choice)
  #get best threshold for test probabilities
  ti = best_threshold(test_probabilities, df_test = df_test)
  print (paste("Best threshold is",ti))
  #get test accuracy
  y_act = df_test$R1
  #print (paste("Length of Test Probabilities",nrow(df_test),length(test_probabilities)))
  return (acc_func(y_act = y_act, pred_probs = test_probabilities, thresh = ti))
}
```

```{r}
train_valid_func(df_train = df_train, df_test = df_valid, k_choice = 20)
```

```{r, include=FALSE}
acc_df2 <- data.frame(matrix(ncol = 0, nrow = length(seq(3,71,2))))
acc_df2$k_kknn = seq(3,71,2)
acc = lapply(seq(3,71,2), function(x) 
  train_valid_func(df_train = df_train, df_test = df_valid, k_choice = x))
acc_df2$acc_validation_set = unlist(acc)
```

```{r}
ggplot(data = acc_df2, aes(x = k_kknn, y = acc_validation_set)) +
  geom_line() +geom_point() +
  geom_vline(xintercept = acc_df2[which.max(acc_df2$acc_validation_set),1],
             color ="magenta", size = 1.5, alpha = 0.5) +
  ggtitle(paste("max validation accuracy for k =",acc_df2[which.max(acc_df2$acc_validation_set),1]))

print (paste("max validation accuracy for k =",acc_df2[which.max(acc_df2$acc_validation_set),1]))
```

rebuild model with optimal k, with

-   Train = Train + Validation

-   Test = Test

```{r}
k_optimal = acc_df2[which.max(acc_df2$acc_validation_set),1]
acc_optimal_k = train_valid_func(df_train = rbind(df_train,df_valid), df_test = df_test, k_choice = k_optimal)
print (paste("With optimal k value, Accuracy on test =",acc_optimal_k))
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```