### 2022-09-06 hw2/anshit/Homework_2.Rmd
added box plots for analysis

---
title: "Homework_2"
output: html_document
date: "2022-09-04"
---

```{r}
# imports
library(ggplot2)
library(cowplot)
library(reshape2)
library(factoextra)
```

## Question 3.1

```{r}
cc_data = read.table('./data 3.1/credit_card_data-headers.txt', header = TRUE, sep='', dec='.')
head(cc_data)
dim(cc_data)
```

## Question 4.2

```{r}
iris_data = read.table('./iris.txt', header = TRUE, sep='', dec='.')
head(iris_data,50)
dim(iris_data)
```

### Exploratory Data Analysis

```{r}
summary(iris_data)
```

```{r}
table(iris_data$Species)
```

```{r}
# Density Plots
my_plots = lapply(colnames(iris_data), function(x){
  p = ggplot(iris_data) + aes_string(x)
  if(x=='Species') {
    p = p + geom_bar()
  }
  else {
    p = p + geom_density()
  }
})

plot_grid(plotlist = my_plots)
```

```{r}
# Correlation Matrix
cor_mat = round(cor(iris_data[,1:4]), 2)
melted_mat = melt(cor_mat)
ggplot(melted_mat, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() + geom_text(aes(Var1, Var2, label=value), color='white', size=3)
```

### Modelling

```{r}
# Calculation of best 'K'
elbow_graph_data <- data.frame(K=integer(), within_sum_of_square=double())
for(k in seq(1, 10, by=1)){
  km.res <- kmeans(as.matrix(scale(iris_data[,1:4])), k, nstart = 25)
  elbow_graph_data[nrow(elbow_graph_data) + 1, ] = c(k, km.res$tot.withinss)
}
```

```{r}
elbow_graph_data
```

```{r}
plot(elbow_graph_data, type='b', ylab='Total Sum of Square Within', main='Elbow Graph', col='blue')
```
```{r}
# Calculation of best 'K'
bw_graph_data <- data.frame(K=integer(), between_sum_of_square=double())
for(k in seq(1, 10, by=1)){
  km.res <- kmeans(as.matrix(scale(iris_data[,1:4])), k, nstart = 25)
  bw_graph_data[nrow(bw_graph_data) + 1, ] = c(k, round(km.res$betweenss,2))
}
#options(scipen=999)
bw_graph_data
```

```{r}
plot(bw_graph_data, type='b', ylab='Betweeen Sum of Square (Intra Cluster)', main='BW SS Graph', col='green')
```

```{r}
# Calculation of best 'K'
ratio_graph_data <- data.frame(K=integer(), ratio=double())
for(k in seq(2, 10, by=1)){
  km.res <- kmeans(as.matrix(scale(iris_data[,1:4])), k, nstart = 25)
  ratio_graph_data[nrow(ratio_graph_data) + 1, ] = c(k, round(km.res$tot.withinss/km.res$betweenss,2))
}
#options(scipen=999)
ratio_graph_data
```
```{r}
plot(ratio_graph_data, type='b', ylab='Ratio Within/Between SS', main='Quality', col='red')
```

Best K is 3 because we are getting marginal benefit after that

```{r}
best_km <- kmeans(as.matrix(scale(iris_data[,1:4])), 3, nstart = 25)
out<-cbind(iris_data,tagged_cluster=best_km$cluster)
out
```
```{r}
boxplot(Sepal.Length~tagged_cluster,data=out)
```

```{r}
out$tagged_cluster<-as.factor(out$tagged_cluster)
ggplot(out, aes(x=tagged_cluster, y=Sepal.Length,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")

out$tagged_cluster<-as.factor(out$tagged_cluster)
ggplot(out, aes(x=tagged_cluster, y=Sepal.Width,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")

out$tagged_cluster<-as.factor(out$tagged_cluster)
ggplot(out, aes(x=tagged_cluster, y=Petal.Length,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")

out$tagged_cluster<-as.factor(out$tagged_cluster)
ggplot(out, aes(x=tagged_cluster, y=Petal.Width,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
```
Petal width and Petal Length are strong indicators for one class (setosa)

```{r}
cm <- table(iris_data$Species, best_km$cluster)
cm
```

```{r}
# function getStatusForClusterAndSpecie(row){
#   if(row$tagged_cluster=="1" & row$Species=="versicolor"){'improper_versicolor'}
#   else if(row$tagged_cluster=="2" & row$Species=="virginica"){'improper_virginica'}
#   else if()
#   
# }
```

```{r}
classifier_df = out[!out$Species=='setosa',]
classifier_df$status = ''
classifier_df$status[(classifier_df$tagged_cluster=="1" & classifier_df$Species=="versicolor")] = 'improper_versicolor'
classifier_df$status[(classifier_df$tagged_cluster=="2" & classifier_df$Species=="virginica")] = 'improper_virginica'
classifier_df$status[(classifier_df$tagged_cluster=="2" & classifier_df$Species=="versicolor")] = 'proper_versicolor'
classifier_df$status[(classifier_df$tagged_cluster=="1" & classifier_df$Species=="virginica")] = 'proper_virginica'

classifier_df

```
```{r}
classifier_df$status<-as.factor(classifier_df$status)
ggplot(classifier_df, aes(x=status, y=Sepal.Length,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")

ggplot(classifier_df, aes(x=status, y=Sepal.Width,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")

ggplot(classifier_df, aes(x=status, y=Petal.Length,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")

ggplot(classifier_df, aes(x=status, y=Petal.Width,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
```

```{r}
out$tagged_cluster<-as.factor(out$tagged_cluster)
ggplot(out, aes(x=tagged_cluster, y=Petal.Width,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
```

```{r}
# renaming predicted clusters according to categorical response variable
# renamed_clusters <- c("versicolor", "setosa", "virginica")[best_km$cluster]
```

```{r}
kmeans_basic_table <- data.frame(best_km$size, best_km$centers)
kmeans_basic_table
```

```{r}
kmeans_basic_df <- data.frame(actual_value=iris_data[,5], prediction=best_km$cluster)
kmeans_basic_df
```

```{r}
plot(scale(iris_data[,1:4]), col=best_km$cluster)
```

```{r}
# prediction_accuracy <- round(sum(kmeans_basic_df$actual_value == kmeans_basic_df$prediction) / nrow(kmeans_basic_df) * 100, 2)
# print(paste('Accuracy Acheived ->', prediction_accuracy, '%' ))
```

### Plotting Clusters

```{r}
fviz_cluster(best_km, data=scale(iris_data[,1:2]), geom=c("point"), ellipse.type = "euclid")
```
***Analysing For K=4***

```{r}
alt_kmeans <- kmeans(as.matrix(scale(iris_data[,1:4])), 4, nstart = 25)
alt_kmeans
```

```{r}
cm <- table(iris_data$Species, alt_kmeans$cluster)
cm
```

```{r}
alt_kmeans$centers
```

```{r}
alt_kmeans_table_size_v_centers <- data.frame(alt_kmeans$size, alt_kmeans$centers)
alt_kmeans_table_size_v_centers
```

```{r}
alt_kmeans_row_pred <- data.frame(actual_value=iris_data[,5], prediction=alt_kmeans$cluster)
alt_kmeans_row_pred
```

```{r}
# clusters = data.frame(alt_kmeans$cluster)
# clusters[clusters$alt_kmeans.cluster==4,]
mapped_data <- data.frame(iris_data[,1:4], prediction=alt_kmeans$cluster)
```

```{r}
# Summary for Cluster 1
summary(mapped_data[mapped_data$prediction==1,1:4])
```

```{r}
# Summary for Cluster 2
summary(mapped_data[mapped_data$prediction==2,1:4])
```

```{r}
# Summary for Cluster 3
summary(mapped_data[mapped_data$prediction==3,1:4])
```

```{r}
# Summary for Cluster 4
summary(mapped_data[mapped_data$prediction==4,1:4])
```

```{r}
plot(scale(iris_data[,1:4]), col=best_km$cluster)
```

```{r}
fviz_cluster(best_km, data=scale(iris_data[,1:2]), geom=c("point"), ellipse.type = "euclid")
```