### 2022-09-12 hw3-1/anshit/HW_3.Rmd
Added HW3 - Initial Commit

---
name: "HW 3 Assignment"
author: "Anshit Verma"
date: "12/09/2022"
output: "html_document"
---

### Question 6.2

```{r}
data = read.table('./temps.txt', sep="\t", dec=".", header = TRUE)
data
```

```{r}
summary(data$X1996)
sd(data$X1996)
```

```{r}
mean = 83.72
st = 0
x = 8
st_df = data.frame(i=double(), st=double())
for(i in 1:nrow(data)){
  st = max(0, st + (mean - data$X1996[i] - x))
  st_df[nrow(st_df)+1,] = c(i, st)
}
plot(st_df, xlab='Day', ylab='St', type='o')
```

```{r}
yearly_mean = data.frame(DAY=data[,1], YearlyMean=rowMeans(data[,-1]))
yearly_mean
```

```{r}
summary(yearly_mean$YearlyMean)
```

```{r}
mean = mean(yearly_mean$YearlyMean)
st = 0
x = 8
st_df = data.frame(i=double(), st=double())
for(i in 1:nrow(yearly_mean)){
  st = max(0, st + (mean - yearly_mean$YearlyMean[i] - x))
  st_df[nrow(st_df)+1,] = c(i, st)
}
plot(st_df, xlab='Day', ylab='St', type='o')
```

# Inferences - 
1. Comparing the above plot with 1996, days when the temp is rising has shifted towards later part of the year, it suggests a change in season cycle.
2. Our max value of our running sum has gotten lower which also suggests warm season in later year.
3. Range of temp values has decreased (min has inc by 8 units) which has shifted temp towards mean.

# TODO - 
1. C and T hyperparameter - how to decide on values.
2. Dividing in buckets to see better changes in YoY temps.
Reference - https://rpubs.com/ssufian/658130