### 2022-09-12 hw3-1/ashish/hw3_test.Rmd
ad sample

---
title: "HW3, Ashish EDA"
output: html_notebook
---

```{r}
data = read.table('../temps.txt', sep="\t", dec=".", header = TRUE)
data
```

```{r}
data_celsius = 
```

```{r}
CUSUM_ad = function(x,days = data$DAY ,C = 0.30,T = 4){
  mean_x = mean(x)
  sd_x = sd(x)
  C = C * sd_x
  #apply cusum
  x1 = lapply(x, function(xi)  (mean_x - xi - C))
  st_df = data.frame(st=double())
  st = 0
  for(i in 1:length(x)){
    #print (paste(i,class(st),class(x1[[i]])))
    st = max(0, st + x1[[i]])
    st_df[i,] = c(st)
  }
  #Find index where St > T
  st_df$Day = days
  st_df["anamoly"] = 0
  st_df$anamoly[st_df$st > T] = 1
  st_df$xt = x
  #plot x with T and if decrease see
  st_df
}
```

```{r}
check = CUSUM_ad(data$X1996)
#library(ggplot2)
par(mfrow=c(2,1))
plot(check$x)
plot(check$st)
#ggplot(check, aes(x = Day, y = st)) + geom_point() + geom_line() +
#  theme(axis.text.x = element_text(angle = 90))
```

```{r}

```

```{r}

```

```{r}

```