### 2022-09-21 hw4-1/anshit/anshit.Rmd

---
title: "Homework-4"
author: "Anshit Verma"
date: "28-09-2022"
---

## Question 8.1

### Example: Real Estate Price

Linear Regression models find use case in the real estate industry to predict selling price for a particular property depending on the prediction variables.

**Objective Function (Y)**: To predict the selling price of a preperty based in the predicting variables for that property
**Predicting Variables (X_i)**: Some of the key predicting variables are given below -

- ***Location***: The location of the property is an important predictor for predicting the selling price. As properties in a good neighborhood will fetch higher prices than those in a relatively worse neighborhoods.

- ***Lot Size***: Lot size is defined as the area of the property. A large lot size is worth more than a smaller lot.

- ***Age of the House (Condition)***: Age of the house is the number of years since the construction of the house. It can also be considered as the condition metric of the house. A house with good condition grade can worth more than one with lower condition grade as it might incure lower maintenance cost.

- ***Garage Capacity***: Capacity of garage is defines as the number of cars a garage can hold. The ability of the garage to hold more cars is an important factor which appreciates the value of a property