### 2022-09-28 hw5-1/ashish/a.0_random_forest.Rmd
submission RF v0

---
title: "Random Forest | Ques 10.1"
author: "Ashish Dhiman"
date: "2022-09-26"
output: html_document
---

```{bash}
ls ../
```

```{r, warning=FALSE}
library(randomForest)
```

```{r}
crime_df = read.table("../uscrime.txt", sep="\t", header= TRUE)
dim(crime_df)
```

#### Data and variable meaning:

Criminologists are interested in the effect of punishment regimes on crime rates. This has been studied using aggregate data on 47 states of the USA for 1960.

| Variable | Description                                                                            |
|-------------------|-----------------------------------------------------|
| M        | percentage of males aged 14--24 in total state population                              |
| So       | indicator variable for a southern state                                                |
| Ed       | mean years of schooling of the population aged 25 years or over                        |
| Po1      | per capita expenditure on police protection in 1960                                    |
| Po2      | per capita expenditure on police protection in 1959                                    |
| LF       | labour force participation rate of civilian urban males in the age-group 14-24         |
| M.F      | number of males per 100 females                                                        |
| Pop      | state population in 1960 in hundred thousands                                          |
| NW       | percentage of nonwhites in the population                                              |
| U1       | unemployment rate of urban males 14--24                                                |
| U2       | unemployment rate of urban males 35--39                                                |
| Wealth   | wealth: median value of transferable assets or family income                           |
| Ineq     | income inequality: percentage of families earning below half the median income         |
| Prob     | probability of imprisonment: ratio of number of commitments to number of offenses      |
| Time     | average time in months served by offenders in state prisons before their first release |
| Crime    | crime rate: number of offenses per 100,000 population in 1960                          |

#Vanilla RF model

```{r}
lapply(crime_df, class)
```

```{r}
crime_df$So <- factor(crime_df$So)
```

```{r}
rf_model_v0 = randomForest(Crime ~ ., data = crime_df)
summary(rf_model_v0)
```

```{r}
rf_model_v0
```

```{r}
rmse_func = function(true,predicted){
  (mean((true-predicted)^2))**0.5
}

rmse_func(true = crime_df$Crime, predicted=rf_model_v0$predicted)
```

```{r}
rf_model_v1 = randomForest(Crime ~ ., data = crime_df, ntree = 100,importance=TRUE)
rmse_func(true = crime_df$Crime, predicted=rf_model_v1$predicted)
```

```{r}
rf_model_v2 = randomForest(log(Crime) ~ ., data = crime_df, ntree = 100,importance=TRUE)
rmse_func(true = crime_df$Crime, predicted=exp(rf_model_v2$predicted))
```

```{r}
varImpPlot(rf_model_v1,sort = TRUE)
```

```{r}
rf_model_v3 = randomForest(Crime ~ . - Ineq -So -Pop, data = crime_df, ntree = 100)
rmse_func(true = crime_df$Crime, predicted=rf_model_v3$predicted)
```

```{r}
rmse_list = sapply(seq(100,1000,20), function(x)  rmse_func(true = crime_df$Crime, predicted=randomForest(Crime ~ ., data = crime_df, ntree = x)$predicted))
plot(x=seq(100,1000,20), y=rmse_list, main = "MSE vs ntrees")
```

```{r}
rmse_list = sapply(seq(2,8,1), function(x)  rmse_func(true = crime_df$Crime, predicted=randomForest(Crime ~ ., data = crime_df, mtry = x)$predicted))
plot(x=seq(2,8,1), y=rmse_list,main="RMSE vs varaibles split")
```

```{r}
rmse_list = sapply(seq(200,1000,20), function(x)  rmse_func(true = crime_df$Crime, predicted=randomForest(Crime ~ ., data = crime_df, ntree = x, mtry=3)$predicted))
plot(x=seq(200,1000,20), y=rmse_list, main = "MSE vs ntrees")
```

```{r}
rf_model_v3 = randomForest(Crime ~ ., data = crime_df, ntree = 250, mtry=3, proximity = TRUE)
rf_model_v3
```

```{r}
prox1=rf_model_v3$proximity
```

### CV

```{r}
n_folds = 5
folds <- sample(rep(1:n_folds, length.out = nrow(crime_df)), size = nrow(crime_df), replace = F)

table(folds)
```

```{r}
rmse_func = function(true,predicted){
  return (mean((true-predicted)^2)**0.5)
}

CV_rf_func = function(test_fold, ntree1, mtry1){
  model <- randomForest(Crime ~ ., data = crime_df[folds != test_fold,],
                        ntree = ntree1, mtry=mtry1)
  #print(model$ntree)
  preds <- predict(model,  crime_df[folds == test_fold,])
  return(rmse_func(true = crime_df[folds == test_fold,"Crime"],predicted = preds))
}
```

```{r}
rmse_list=c()

for (ntree_i in seq(10,400,50)) {
  for (mtry_i in seq(2,15,1)) {
    #RMSE on each CV Fold for ntree_i,mtry_i
    rmse_cv = mean(unlist(lapply(seq(1:5), function(fold) 
      CV_rf_func(test_fold = fold, ntree1 = ntree_i, mtry1 = mtry_i))))
    print (paste("For tree:",ntree_i,"mtry:",mtry_i,"rmse:",rmse_cv))
    rmse_list = c(rmse_list, rmse_cv)
}
}

rmse_df = cbind.data.frame(seq(10,400,50),seq(2,15,1),rmse_list)
```

```{r}
names(rmse_df) = c("ntree","mtry","rmse")
```

```{r}
rmse_df[which.min(rmse_df$rmse),]
```

```{r}
summary(rmse_df$rmse)
```