### 2022-09-28 hw5-1/submission.Rmd

---
title: "ISYE 6501"
output:
  html_document:
    df_print: paged
  pdf_document: default
date: "2022-09-04"
---

# Submission HW4 \| Fall 22

-   Ashish Dhiman \| [ashish.dhiman\@gatech.edu](mailto:ashish.dhiman@gatech.edu){.email}

-   Abhinav Arun \| [aarun60\@gatech.edu](mailto:aarun60gatech.edu){.email}

-   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}

[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}

## Question 10.1

```{r, warning=FALSE, message=FALSE}
set.seed(77)

library(randomForest)
library(ggplot2)
library(cowplot)
library(ggplot2)
library(reshape2)
library(grid)
library(MASS)
library(rpart)
library(rpart.plot)
```

```{r}
crime_data<-read.table(file = "./uscrime.txt", sep = "\t",header=TRUE)
head(crime_data)
dim(crime_data)
summary(crime_data)
```

### EDA

### Data Distribution and Pairwise Corelation

```{r}
my_plots <- lapply(names(crime_data), function(var_x){
  p <- 
    ggplot(crime_data) +
    aes_string(var_x)

  if(var_x %in% list("So")) {
    p <- p + geom_bar()

  } else {
    p <- p + geom_density()
  }

})

plot_grid(plotlist = my_plots)

### Correlation
cormat <- round(cor(crime_data),2)
cormat[upper.tri(cormat)] <- NA
melted_cormat <- melt(cormat)
# plotting the correlation heatmap

ggplot(data = melted_cormat, aes(x=Var1, y=Var2,
                                   fill=value)) +
geom_tile() +
geom_text(aes(Var1, Var2, label = value),
          color = "white", size = 3)
```

### Decision Tree

#### Training and evaluating on the same training dataset (not a good method for generalizing) , better to do cross validation

```{r}
# Vanilla decision tree with a fixed value of complexity parameter of 0.01 and minsplit of 20
reg_tree<-rpart(Crime~.,data=crime_data,method="anova",control=c(cp=0.01,minsplit=20))
rmse<-sqrt(sum((crime_data$Crime - predict(reg_tree,type=c("vector")))**2)/nrow(crime_data))
rmse
```

```{r}
rpart.plot(reg_tree)
```

[**#Analysis**]{style="color:red"}

-   Observations & Points for further exploration:

-   The above decision tree uses 3 variables NW , Pop, LF and Po1 to create a decision tree. (2 of them Pop and Po1 have a linear correlation with dependent variable i.e. Crime)

-   One food for thought is that this is a regression tree and we are imputing the mean values of all the points that fall in a leaf or terminal node , so incase if we have an outlier , then those will drive the value of mean and hence cause higher RMSE or SSE. So , it would be good to see the effect of removing or capping outliers and then performing decision tree.

-   The smaller the value of complex parameter , the model is expected to have high variance as a split would be performed if we get such a marginal increase in entropy . Therefore , to prevent our model from overfitting amd having high variance , it is better if we not choose too small values of cp i.e. complexity parameter.

-   The default value of max depth parameter here is 30 but given that we have only 47 nodes in our dataset , it dfoes not make much sense to have tree models greater than or equal to 3 or 4 levels of depth

-   minsplit hyperparameter : The default value of this parameter is 20 i.e. a split would be performed if we have atleast 20 datapoints in a node , although it is generally good to have higher values of "minsplit" parameter to avoid overfitting but since in this problem statement , the total \# of datapoints is only 47 , so this is one parameter where we can tweak in a bit .

-   minbucket hyperparameter : Similarly , this hyperparameter means the number of datpoints in a terminal node and thus here also , it makes sense to have a smaller value for this hyperparameter as we do not have large dataset for training our model.

#### Perform hyperparameter tuning minsplit and cp(complexity hyperparameter)

```{r}
folds <- sample(rep(1:5, length.out = nrow(crime_data)), size = nrow(crime_data), replace = F)
pred_table<-data.frame(min_split=numeric(0),cp=double(0),rmse=double(0))
for(j in seq(0.001,0.02,0.001)){
for(i in seq(5,25,2)){
  CV_mse <- lapply(1:5, function(x){ #5 corresponds to the number of folds defined earlier
  model <- rpart(Crime ~ ., data = crime_data[folds != x,],method="anova",control=c(minsplit=i,cp=j))
  preds <- predict(model,  crime_data[folds == x,], type="vector")
  sum_sse<-sum((crime_data[folds==x,"Crime"]-preds)**2)
  rmse<-sqrt(sum_sse/nrow(crime_data[folds==x,]))
  #total_rmse<-total_rmse+rmse
  return(rmse)
   })
pred_table[nrow(pred_table)+1,]=c(i,j,mean(unlist(CV_mse)))
}
}
```

#### Identifying the optimal value of the hyperparameters cp and min_split

```{r}
options(scipen=999)
pred_table[which.min(pred_table$rmse),c("min_split","cp")]
```

-   From the above output , we get the the optimal value of cp as 0.001 and min_split as 15

```{r}
reg_cv_tree<-rpart(Crime~.,data=crime_data,method="anova",control=c(cp=0.001,minsplit=15))
rpart.plot(reg_cv_tree)
```

[**#Analysis**]{style="color:red"}

-   Note: The above set of optimal parameters reproduces the same decision tree that we had done initially with fixed values of these 2 parameters.This can be because of less \# of datapoints. We can try reducing the number of minsplits which will lead to a higher variance model and we can try iterating to get the optimal maxDepth hyperparameter.

```{r}
pred_table<-data.frame(min_split=numeric(0),maxDepth=double(0),rmse=double(0))
for(j in seq(2,15,1)){
for(i in seq(5,25,2)){
  CV_mse <- lapply(1:5, function(x){ #5 corresponds to the number of folds defined earlier
  model <- rpart(Crime ~ ., data = crime_data[folds != x,],method="anova",control=c(minsplit=i,maxdepth=j))
  preds <- predict(model,  crime_data[folds == x,], type="vector")
  sum_sse<-sum((crime_data[folds==x,"Crime"]-preds)**2)
  rmse<-sqrt(sum_sse/nrow(crime_data[folds==x,]))
  #total_rmse<-total_rmse+rmse
  return(rmse)
   })
pred_table[nrow(pred_table)+1,]=c(i,j,mean(unlist(CV_mse)))
}
}  
#rpart.plot(rpart(Crime ~ ., data = crime_data,method="anova",minsplit=10,cp=0.01))
#plot(pred_table)
```

[**#Analysis**]{style="color:red"}

-   Identifying the optimal value of the hyperparameter maxDepth (which comes out to be around 2 with min_split within range of 5-15 (rmse is same for different values of min_split))

```{r}
options(scipen=999)
pred_table[which.min(pred_table$rmse),c("maxDepth")]
```

```{r}
reg_tree_final<-rpart(Crime~.,data=crime_data,method="anova",control=c(cp=0.001,minsplit=15,maxdepth=2))
rmse_final<-sqrt(sum((crime_data$Crime - predict(reg_tree_final,type=c("vector")))**2)/nrow(crime_data))
rmse_final
```

**Conclusion : The RMSE value for the decision tree model using the optimal values of the hyperparameter is `r round(rmse_final,2)`**

### Random Forest

```{r}
crime_df <- crime_data
crime_df$So <- factor(crime_df$So)
```

#### Build Vanilla RF model

```{r}
rf_model_v0 = randomForest(Crime ~ ., data = crime_df, importance=TRUE)
summary(rf_model_v0)

rmse_func = function(true,predicted){
  (mean((true-predicted)^2))**0.5
}

rmse_func(true = crime_df$Crime, predicted=rf_model_v0$predicted)

varImpPlot(rf_model_v0,sort = TRUE)
```

[**#Analysis**]{style="color:red"}

From variable importance we can see that the some variables like, So, U1 and U2 are not very helpful. Hence it might serve us better to drop these variables, bcos they might add more noise to our model instead of actual information capture. In the CV stage below, we will test the models without these variables.

Variable Importance here is given basis two metrics, drop in MSE when the variable is not included in feature set, and second the Information gain (or purity) of a node, if the variable is used to split it.

```{r}
rf_model_v1 = randomForest(Crime ~ . -U1 -M.F -U2, data = crime_df, importance = TRUE)
varImpPlot(rf_model_v1)
```

### Build RF and tune hyperparams with CV

#### Build CV folds

```{r}
n_folds = 5
folds <- sample(rep(1:n_folds, length.out = nrow(crime_df)), size = nrow(crime_df), replace = F)

table(folds)
```

[**#Analysis**]{style="color:red"}

We have created uniform CV folds here.

#### Function for RF on CV fold

```{r}
CV_rf_func = function(test_fold, ntree1, mtry1){
  model1 <- randomForest(Crime ~ ., data = crime_df[folds != test_fold,],
                        ntree = ntree1, mtry=mtry1)
  model2 <- randomForest(Crime ~ . -U1 -M.F -U2, data = crime_df[folds != test_fold,],
                        ntree = ntree1, mtry=mtry1)
  model3 <- randomForest(Crime ~ . -U1 -U2 -Time, data = crime_df[folds != test_fold,],
                        ntree = ntree1, mtry=mtry1)

  preds1 <- predict(model1,  crime_df[folds == test_fold,])
  preds2 <- predict(model2,  crime_df[folds == test_fold,])
  preds3 <- predict(model3,  crime_df[folds == test_fold,])

  rmse1 = rmse_func(true = crime_df[folds == test_fold,"Crime"],predicted = preds1)
  rmse2 = rmse_func(true = crime_df[folds == test_fold,"Crime"],predicted = preds2)
  rmse3 = rmse_func(true = crime_df[folds == test_fold,"Crime"],predicted = preds3)

  return(c(rmse1,rmse2,rmse3))
}

```

```{r, warning=FALSE, message=FALSE}
rmse_list1=c()
rmse_list2=c()
rmse_list3=c()
ntree_list=c()
mtry_list=c()

for (ntree_i in seq(30,130,10)) {
  for (mtry_i in seq(2,6,1)) {
    #RMSE on each CV Fold for ntree_i,mtry_i
    rmse_cv_all = lapply(seq(1:5), function(fold) 
      CV_rf_func(test_fold = fold, ntree1 = ntree_i, mtry1 = mtry_i))
    rmse_cv_df = do.call(rbind.data.frame, rmse_cv_all)
    rmse_cv1 = mean(unlist(rmse_cv_df[,1]))
    rmse_cv2 = mean(unlist(rmse_cv_df[,2]))
    rmse_cv3 = mean(unlist(rmse_cv_df[,3]))
    #print (paste("For tree:",ntree_i,"mtry:",mtry_i))
    rmse_list1 = c(rmse_list1, rmse_cv1)
    rmse_list2 = c(rmse_list2, rmse_cv2)
    rmse_list3 = c(rmse_list3, rmse_cv3)
    ntree_list = c(ntree_list, ntree_i)
    mtry_list = c(mtry_list, mtry_i)
}
}
```

```{r}
rmse_df = cbind.data.frame(ntree_list,mtry_list,
                           rmse_list1,rmse_list2,rmse_list3)
names(rmse_df)=c("ntree","mtry","rmse_model1","rmse_model2","rmse_model3")
```

```{r}
rmse_df
```

### Let us study the impact of these two hyper-parameters on RMSE

```{r}
ggplot(rmse_df[rmse_df$mtry<4,], aes(x = ntree, y = rmse_model1, group = mtry, color = mtry)) +
  geom_point() + geom_line()

ggplot(rmse_df[rmse_df$mtry<4,], aes(x = ntree, y = rmse_model2, group = mtry, color = mtry)) +
  geom_point() + geom_line()

ggplot(rmse_df[rmse_df$mtry<4,], aes(x = ntree, y = rmse_model3, group = mtry, color = mtry)) +
  geom_point() + geom_line()
```

[**#Analysis**]{style="color:red"}

Because our data is so small; the effect of ntree and mtry is very sporadic, with high interaction between them. Also the trend varies as per the choice of the model(i.e the choice of predictors).

Ideally we would have expected the RMSE to exhibit a 'U' shaped trend with out hyper parameters, where adding complexity to model (i.e. more num of trees to average upon or more features to split nodes upon) first decreases RMSE, but after a point it starts overfitting, and RMSE starts increasing again.

```{r}
print ("Min RMSE model 1")
rmse_df[which.min(rmse_df[,"rmse_model1"]),]

print ("Min RMSE model 2")
rmse_df[which.min(rmse_df[,"rmse_model2"]),]

print ("Min RMSE model 3")
rmse_df[which.min(rmse_df[,"rmse_model3"]),]
```

### Best Model is therefore given for:

1.  Dropping variables: U1, M.F and U2
2.  number of trees = 40
3.  mtry = 4

Not only thus by RMSE, but by intution too this is a simpler model since it has lesser features, lower number of trees, as well as lower number of features to split on in RF.

Using the above hyper-parameters we can build the best model on full data.

```{r}
rf_final_model <- randomForest(Crime ~ . -U1 -M.F -U2, data = crime_df, ntree = 40, mtry=4, importance = TRUE)
rmse_func(true=crime_df$Crime,predicted = rf_final_model$predicted)
```

```{r}
varImpPlot(rf_final_model)
```

[**#Analysis**]{style="color:red"}

From the variable importance plot we see, that the following variable feature very highly in Importance:

1.  Po2
2.  NW
3.  Prob

### Comparison between Decision Tree and Random Forest

[**#Analysis**]{style="color:red"}

Note: while the above 3 variables are also significant features here is the variable importance of variable "Pop". While this variable is very important for Decision trees, it is not that important for Random Forest. One possible reason for this could be the notion of randomness involved in RF.

Another point to note here is that after Hyperparameter tuning, Decision tree is giving lower RMSE, and is most likely over-fitting the data-set, in comparison to Random Forest, and if we had a test data, RF model would most likely perform better, because of better generalization.

## Question 10.2

## Question 10.3

```{r, include=FALSE}
library(caret)
library(ROCR)
```

```{r}
german_data = read.table("./germancredit.txt")
german_data
```

```{r}
cols = colnames(german_data)
for(i in seq(1, length(cols))){
  if(class(german_data[, cols[i]]) == "character"){
    german_data[, cols[i]]= as.factor(german_data[, cols[i]])
  }
}
```

### Exploratory Data Analysis

```{r}
summary(german_data)
```

We can see that the data is uniformly distributed, both for the numerical features as well as categorical features.

```{r}
par(mfrow=(c(1,2)))
plot(german_data$V5)
plot(german_data$V2)
```

2 features have suspiciously large max values (comparing to the 3rd quartile).

```{r}
# (1 = Good, 2 = Bad) -> (0 = Good, 1 = Bad)
german_data['V21'] = lapply(german_data['V21'], function(x){
  x-1
})
german_data[,"V21"] = as.factor(german_data[,"V21"])
german_data
```

```{r}
split_data = function(dataset){
  ## Sampling train and test data

  ## 75% of the sample size
  smp_size <- floor(0.75 * nrow(dataset))

  ## set the seed to make your partition reproducible
  set.seed(123)
  train_ind <- sample(seq_len(nrow(dataset)), size = smp_size)

  train <- dataset[train_ind, ]
  test <- dataset[-train_ind, ]
  return(list("train"=train, "test"=test))
}
```

### Splitting the data into train and test

```{r}
df_list = split_data(german_data)
train = df_list$train
test = df_list$test
summary(test$V21)
```

### Training vanilla model with all features

```{r}
model <- glm(V21 ~.,family=binomial(link='logit'),data=train)
model
```

```{r}
summary(model)
```

```{r}
get_roc_plot_and_auc = function(model, test){
  p <- predict(model, newdata=test[, 1:(length(test)-1)], type="response")
  pr <- prediction(p, test$V21)
  prf <- performance(pr, measure = "tpr", x.measure = "fpr")
  auc <- performance(pr, measure = "auc")
  auc <- auc@y.values[[1]]
  return (list("prf"=prf, "auc"=auc, "prediction"=p))
}
```

### Plotting the ROC Curve and calculating AUC

```{r}
roc = get_roc_plot_and_auc(model, test)
plot(roc$prf)
roc$auc
```

[**#Analysis**]{style="color:red"}

As we can see that AUC value is 0.7846446 that is close to 1. Hence our model is above average in being able to separate between classes.

```{r}
predicted = ifelse(roc$prediction > 0.5,1,0)
cor = confusionMatrix(data=as.factor(predicted), reference=as.factor(test$V21), positive="1")
cor
```

```{r}
cor$byClass
```

[**#Analysis**]{style="color:red"}

With and initial (default) threshold of 0.5, we can see that our model achieves an accuracy of 75.6% with a recall of 54.3% and precision of 64.7%. We can try dropping some statistically insignificant features to better improve the metrics of our model.

### Dropping Features

```{r}
# We can drop statistically insignificant features
drop.features = german_data
drops <- c("V7","V17", "V11", "V13", "V16", "V12")
drop.features = drop.features[ , !(names(drop.features) %in% drops)]
```

```{r}
df.list.1 = split_data(drop.features)
train.1 = df.list.1$train
test.1 = df.list.1$test
```

```{r}
model.1 <- glm(V21 ~.,family=binomial(link='logit'),data=train.1)
model.1
```

```{r}
summary(model.1)
```

```{r}
roc.1 = get_roc_plot_and_auc(model.1, test.1)
plot(roc.1$prf)
roc.1$auc
```

[**#Analysis**]{style="color:red"}

We can observe that dropping statistically insignificant features help us achieve a higher AUC value. Hence this increases our model's ability to separate between the classes.

```{r}
predicted.1 = ifelse(roc.1$prediction > 0.5,1,0)
cor.1 = confusionMatrix(data=as.factor(predicted.1), reference=as.factor(test.1$V21), positive="1")
cor.1
```

```{r}
cor.1$byClass
```

[**#Analysis**]{style="color:red"}

Again we can observe that with the initial (default) value of threshold - 0.5, we have increased the accuracy of our model. While the recall remains unchanged, we were able to increase the precision.

### Determining Threshold by F-Beta Value

```{r}
get_recall_and_precision = function(cor){
  cor_table = data.frame(cor$table)
  tn = cor_table[cor_table$Prediction==0 & cor_table$Reference==0, "Freq"]
  fn = cor_table[cor_table$Prediction==0 & cor_table$Reference==1, "Freq"]
  tp = cor_table[cor_table$Prediction==1 & cor_table$Reference==1, "Freq"]
  fp = cor_table[cor_table$Prediction==1 & cor_table$Reference==0, "Freq"]
  recall = tp / (tp + fn)
  precision = tp / (tp + fp)
  return (c("recall"=recall, "precision"=precision))
}
```

[**#Analysis**]{style="color:red"}

We can use the F-Beta value for determining the best value of threshold. As we are given that a classifying a "bad" customer as "good" is 5 times worse than classifying "good" customer as "bad". We need to determine the best value of threshold that will help us increase recall and accordingly lower the value of false negatives (positive class is 1 i.e. "bad").

```{r}
rp.df = data.frame("threshold"=double(), "recall"=double(), "precision"=double(), f_beta=double())
for(i in seq(0, 100, 1)){
  beta=2
  predicted.2 = ifelse(roc.1$prediction > i/100,1,0)
  cor.2 = confusionMatrix(data=as.factor(predicted.2), reference=as.factor(test.1$V21), positive = "1")
  rp = get_recall_and_precision(cor.2)
  f_beta = ((1 + beta^2) * rp[2] * rp[1]) / (beta^2 * rp[2] + rp[1])
  rp.df[nrow(rp.df)+1,] = c("threshold"=i/100, rp, f_beta)
}
```

[**#Analysis**]{style="color:red"}

We are using F2 specifically which helps us weight recall more than precision. Since we want to maximize recall and minimize false negative due to higher cost associated with it, F2 will help us determine the best threshold value. Since F2 can range from 0 to 2, the higher the value of F2 - the better the threshold to minimize false negative.

```{r}
rp.df
```

```{r}
# Max value of F Beta Score
rp.df[which.max(rp.df$f_beta),]
```

***Plotting the best value of threshold for maximum F2 value***

```{r}
plot(x=rp.df$threshold, y=rp.df$f_beta, type="o", xlab="Probability Threshold", ylab="F Beta Score", main="F2 v/s Threshold")
points(x=rp.df$threshold[which.max(rp.df$f_beta)], y=rp.df$f_beta[which.max(rp.df$f_beta)], pch = 18, col = "red", type = "b", lty = 2)
```

[**#Analysis**]{style="color:red"}

We can see from the above plot that best value that minimizes false negatives is 0.11.

### Determining Threshold by Calculating Error Rate

```{r}
get_cm_value = function(cor){
  cor_table = data.frame(cor.2$table)
  tn = cor_table[cor_table$Prediction==0 & cor_table$Reference==0, "Freq"]
  fn = cor_table[cor_table$Prediction==0 & cor_table$Reference==1, "Freq"]
  tp = cor_table[cor_table$Prediction==1 & cor_table$Reference==1, "Freq"]
  fp = cor_table[cor_table$Prediction==1 & cor_table$Reference==0, "Freq"]
  return(list(tn=tn,fn=fn,tp=tp,fp=fp))
}
```

[**#Analysis**]{style="color:red"}

Another way determine the threshold is by calculating the error value by weighing the false negatives more than false positive. As we know that false negatives are 5 time more costly than false positive, we can use that information while calculating our error value.
We need to choose the value of threshold which minimizes the error.

```{r}
## Minimizing the error
## Since we know FN is 5 times worse than FP, we minimize the error that gives us the best threshold

error.df = data.frame("threshold"=double(), "error"=double())
for(i in seq(0, 100, 1)){
  beta=2
  predicted.2 = ifelse(roc.1$prediction > i/100,1,0)
  cor.2 = confusionMatrix(data=as.factor(predicted.2), reference=as.factor(test.1$V21), positive = "1")
  cm.values = get_cm_value(cor.2)
  total_error = (5*cm.values$fn + cm.values$fp)/nrow(german_data)
  error.df[nrow(error.df)+1,] = c("threshold"=i/100, "error"=total_error)
}
```

```{r}
error.df[which.min(error.df$error),]
```

[**#Analysis**]{style="color:red"}

We can again see that our error value is minimum at the threshold value of 0.11. This is in sync with our value of threshold which we determined via F-beta method. This this value of threshold is optimal.

```{r}
plot(error.df)
points(x=error.df$threshold[which.min(error.df$error)], y=error.df$error[which.min(error.df$error)], pch = 18, col = "red", type = "b", lty = 2)
```

### Final Model

```{r}
final.predicted = ifelse(roc.1$prediction > 0.11,1,0)
final.cor = confusionMatrix(data=as.factor(final.predicted), reference=as.factor(test.1$V21), positive="1")
final.cor
```

```{r}
final.cor$byClass
```

[**#Analysis**]{style="color:red"}

By choosing our threshold to be 0.11, we can achieve a recall value of 0.91. While our precision has gone down, we are not making mistakes with false negatives which are more costly to us.

In our test data, the no. of "bad" classified data points was ~33%. We were able to achieve a threshold value of 0.11 which is 1/3rd of that value.