### 2022-10-11 hw7/ashish/lasso.Rmd
hw 7: lasso

```{bash}
ls ../
```

```{r, warning=FALSE}
library(glmnet)
```

```{r}
crime_df = read.table("../uscrime.txt", sep="\t", header= TRUE)
dim(crime_df)
```

```{r}
names(crime_df)
```

```{r}
x = as.matrix(crime_df[,names(crime_df) != "Crime"])
y = as.matrix(crime_df[,"Crime"])
dim(x)
dim(y)
```

```{r}
fit <- glmnet(x, y, alpha = 1, nlambda = 20, standardize = TRUE)
```

```{r}
print(fit)
```

```{r}
plot(fit, xvar = "lambda", label = TRUE)
```

```{r}
cvfit <- cv.glmnet(x = x, y = y, alpha = 1, nlambda = 30, standardize = TRUE, type.measure = "mse", nfolds = 5)
```

Because we have only 47 points in the data, we don't want fit too many models during the CV stage

```{r}
plot(cvfit)
```

Above plot gives CV error against values of lambda. With increasing Lambda the weight of regularization factor increases, and therefore the model is encouraged, to select lower number of predictors.

```{r}
print (cvfit)
```

Two special values along the 𝜆λ sequence are indicated by the vertical dotted lines. `lambda.min` is the value of λ that gives minimum mean cross-validated error, while `lambda.1se` is the value of λ that gives the most regularized model such that the cross-validated error is within one standard error of the minimum.

```{r}
coef(cvfit, s = "lambda.min")
```

```{r}
coef(cvfit, s = "lambda.1se")
```

```{r}
cvfit$lambda.1se
```

```{r}
fit_1se <- glmnet(x, y, alpha = 1, lambda = cvfit$lambda.1se, standardize = TRUE)
coef(fit_1se)

fit_min <- glmnet(x, y, alpha = 1, lambda = cvfit$lambda.min, standardize = TRUE)
coef(fit_min)
```

```{r}
predict_min = predict(fit_min, newx = x, s = "lambda.min")
predict_1se = predict(fit_1se, newx = x, s = "lambda.1se")
```

```{r}
rmse_func = function(true,predicted){
  (mean((true-predicted)^2))**0.5
}
```

### RMSE for

```{r}
rmse_func(true=y,predicted = predict_min)
```

```{r}
rmse_func(true=y,predicted = predict_1se)
```